A memory frontier for complex synapses: Supplementary material

### 2013-05-30 NIPS/SynapseNIPSsupp.tex
almost final version

@@

Consider the first time the process leaves state $i$.
On average, this will take time $\Lb_{ii}$.
With probability $\Pb_{ij}$, it will go directly to $j$, so the conditional mean time would be $\Lb_{ii}$.
On the other hand, if it goes to some other state, $k$, with probability $\Pb_{ik}$, the conditional mean time would be $\fptb_{kj}+\Lb_{ii}$.
Combining these, we get the recursion relation
%
\begin{equation}\label{eq:fptrecursion}
  \begin{aligned}
  \fpt_{ij} &= \sum_{k\neq j} \Pb_{ik}(\fptb_{kj}+\Lb_{ii}) + \Pb_{ij}\Lb_{ii} \\
    &= \sum_{k\neq j} \Pb_{ik}\fptb_{kj} + \sum_{k} \Pb_{ik}\Lb_{ii}\\
    &= \sum_{k\neq j} \Pb_{ik}\fptb_{kj} + \Lb_{ii},\\
   \fpt &= \Pb \fptb + \Lb\onem,
  \end{aligned}
\end{equation}
%
or, using \eqref{eq:defDLP},
%
\begin{equation}\label{eq:fptrecdg}
  \fpt\dg = \Lb\MM\fptb+\Lb\onem.
\end{equation}
%

\begin{thm}
  The recurrence times are given by
  %
  \begin{equation}\label{eq:recurtime}
    \fpt\dg = \Lb\D.
  \end{equation}
  %
\end{thm}
\begin{proof}
  Pre-multiply \eqref{eq:fptrecdg} by $\eq\Lb\inv$:
  %
  \begin{equation*}
    \eq\Lb\inv\fpt\dg = \onev\trans,
  \end{equation*}
  %
  or in component form
  %
  \begin{equation*}
    \eq_i \Lb_{ii}\inv \fpt\dg_{ii} = 1.
  \end{equation*}
  %
  This is the same as \eqref{eq:recurtime} in component form: $\fpt\dg_{ii} = \Lb_{ii}/\eq_i$.
\end{proof}
I think the extra factor of $\Lb_{ii}$, compared to the discrete case \cite[Th.4.4.5]{kemeny1960finite}, occurs because in this case we are demanding that the process leaves the initial state once before returning, whereas in the discrete case we only measure the time it takes to go to the initial state after the first time-step.

We can substitute this into \eqref{eq:fptrecdg} to get a recursion relation for the off diagonal part:
%
\begin{equation}\label{eq:fptbrec}
  \MM\fptb = \D-\onem.
\end{equation}
%
\begin{thm}
  The off-diagonal solution of \eqref{eq:fptbrec} is unique.
\end{thm}
\begin{proof}
  Suppose we have two solutions for the off-diagonal part, $\fptb$ and $\fptb'$.
  Subtracting \eqref{eq:fptbrec} for these two solutions gives
  %
  \begin{equation*}
    \MM(\fptb-\fptb')=0.
  \end{equation*}
  %
  As we are assuming an ergodic process, the zero eigenvalue is non-degenerate.
  Therefore
  %
  \begin{equation*}
    \fptb-\fptb' = \onev\mathbf{y},
  \end{equation*}
  %
  for some row-vector $\mathbf{y}$. Looking at the diagonal components
  %
  \begin{equation*}
    \mathbf{y}_{i} = \fptb_{ii}-\fptb'_{ii} = 0,
  \end{equation*}
  %
  and therefore $\fptb-\fptb'=0$.
\end{proof}

\begin{thm}
  The off-diagonal mean first passage times are given by
  %
  \begin{equation}\label{eq:fptfund}
    \fptb = (\onem\fund\dg - \fund)\D.
  \end{equation}
  %
\end{thm}
\begin{proof}
  We just need to show that this satisfies \eqref{eq:fptbrec} and that the diagonal elements vanish.
  Premultiply \eqref{eq:fptfund} by $\MM$:
  %
  \begin{equation*}
  \begin{aligned}
    \MM\fptb &= -\MM\fund\D\\
     &= (\I-\onev\eq)\D \\
     &= \D-\onem,
  \end{aligned}
  \end{equation*}
  %
  where we used \eqref{eq:fundQZ} and \eqref{eq:pdotD}.
  The vanishing of the diagonal elements follows trivially from the component form of \eqref{eq:fptfund}:
%
\begin{equation}\label{eq:fptfundcmpt}
  \fptb_{ij} = \frac{\fund_{jj}-\fund_{ij}}{\eq_j}.
\end{equation}
%

\end{proof}

@@

One can show that the equilibrium probabilities, $\eq$, are invariant under these operators:
%
\begin{equation}\label{eq:sareacoeffincrprob}
  \Delta_{gh} \eq_i = 0,
\end{equation}
%
which makes it possible to integrate the perturbation:
%
\begin{equation}\label{eq:areacoeffincrfinite}
  \W\potdep  \to \W\potdep  + \D\boldsymbol{\epsilon},
  \qquad
  \begin{aligned}
  \boldsymbol{\epsilon} &= \boldsymbol{\epsilon}\trans,
  \\
  \boldsymbol{\epsilon} \onev &= 0.
  \end{aligned}
\end{equation}
%
But more interestingly:
%
\begin{align}
%  \begin{aligned}
\label{eq:areacoeffincr}
    \Delta_{gh}\CL &=
      \frac{\mu\pot _{gh}+\mu\dep _{gh}}{\eq_g} + \frac{\mu\pot _{hg}+\mu\dep _{hg}}{\eq_h}
      + 2r\prn{c_g-c_h}\prn{\eta^+_{g} - \eta^+_{h}},\\
%  \end{aligned}
\end{align}
%
where $c_k$ were defined in \eqref{eq:areacoeffs}.

@@

One can show that the equilibrium probabilities, $\eq$, are invariant under these operators:
%
\begin{equation}\label{eq:shortcutprob}
  \widetilde{\Delta}\potdep _{g,m} \eq_i = 0,
\end{equation}
%
which makes it possible to integrate the perturbation:
%
\begin{equation}\label{eq:shortcutfinite}
  \W\potdep  \to \W\potdep  + \D\boldsymbol{\epsilon}^{\pm(g,m)},
  \qquad
  \begin{aligned}
    &\prn{\boldsymbol{\epsilon}^{\pm(g,m)}}_{g,g\pm m}
      &\!\!\!\!\!\!&= -\epsilon,\\
    &\prn{\boldsymbol{\epsilon}^{\pm(g,m)}}_{g\pm k,g\pm(k+1)}
      &\!\!\!\!\!\!&= \epsilon
        &\;
        &\forall\, k \in [0,m-1],\\
    &\prn{\boldsymbol{\epsilon}^{\pm(g,m)}}_{g\pm k,g\pm k}
      &\!\!\!\!\!\!&= -\epsilon
        &
        &\forall\, k \in [1,m-1].
  \end{aligned}
\end{equation}
%
But more interestingly for our purposes:
%
\begin{align}
%  \begin{aligned}
\label{eq:shortcutarea}
    \widetilde{\Delta}\potdep _{g,m}\CL &=
      \brk{ \sum_{k=0}^{m-1} \frac{\mu\potdep _{g\pm k,g\pm(k+1)}}{\eq_{g\pm k}}
      - \frac{\mu\potdep _{g,g\pm m}}{\eq_g}}
      + 2rf\potdep  \sum_{k=0}^{m-1} \prn{\eta^+_{g\pm k} - \eta^+_{g\pm(k+1)}} \prn{c_{g\pm k} - c_{g}},
%  \end{aligned}
\end{align}
%

>>>>

Although the theory and derivations outlined in the main text are relatively self-contained, here we provide more details for the aid of the reviewer.

@@

This matrix is given by
%
\begin{equation}\label{eq:fpt}
  \fpt=(\onem\fund\dg - \fund + \Lb)\D,
\end{equation}
%
see \cite{Yao1985fpt} for a proof.
We can separate this into its diagonal and off-diagonal pieces.

The recurrence times are given by
%
\begin{equation}\label{eq:recurtime}
  \fpt\dg = \Lb\D.
\end{equation}
%
or in component form
%
\begin{equation*}
  \eq_i \Lb_{ii}\inv \fpt\dg_{ii} = 1.
\end{equation*}
%
The extra factor of $\Lb_{ii}$, compared to the discrete case \cite[Th.4.4.5]{kemeny1960finite}, occurs because in this case we are demanding that the process leaves the initial state once before returning, whereas in the discrete case we only measure the time it takes to go to the initial state after the first time-step.

The off-diagonal mean first passage times are given by
%
\begin{equation}\label{eq:fptfund}
  \fptb = (\onem\fund\dg - \fund)\D.
\end{equation}
%
or in component form of \eqref{eq:fptfund}:
%
\begin{equation}\label{eq:fptfundcmpt}
  \fptb_{ij} = \frac{\fund_{jj}-\fund_{ij}}{\eq_j}.
\end{equation}
%

@@

One can show that the equilibrium probabilities, $\eq$, are invariant under these operators using \eqref{eq:diffpT}:
%
\begin{equation}\label{eq:sareacoeffincrprob}
  \Delta_{gh} \eq_i = 0,
\end{equation}
%
which makes it possible to integrate the perturbation:
%
\begin{equation}\label{eq:areacoeffincrfinite}
  \W\potdep  \to \W\potdep  + \D\boldsymbol{\epsilon},
  \qquad
  \begin{aligned}
  \boldsymbol{\epsilon} &= \boldsymbol{\epsilon}\trans,
  \\
  \boldsymbol{\epsilon} \onev &= 0.
  \end{aligned}
\end{equation}
%
But more interestingly:
%
\begin{align}
%  \begin{aligned}
\label{eq:areacoeffincr}
    \Delta_{gh}\CL &=
      \frac{\mu\pot _{gh}+\mu\dep _{gh}}{\eq_g} + \frac{\mu\pot _{hg}+\mu\dep _{hg}}{\eq_h}
      + 2r\prn{c_g-c_h}\prn{\eta^+_{g} - \eta^+_{h}},\\
%  \end{aligned}
\end{align}
%
where $c_k$ were defined in \eqref{eq:areacoeffs}.

@@

One can show that the equilibrium probabilities, $\eq$, are invariant under these operators \eqref{eq:diffpT}:
%
\begin{equation}\label{eq:shortcutprob}
  \widetilde{\Delta}\potdep _{g,m} \eq_i = 0,
\end{equation}
%
which makes it possible to integrate the perturbation:
%
\begin{equation}\label{eq:shortcutfinite}
  \W\potdep  \to \W\potdep  + \D\boldsymbol{\epsilon}^{\pm(g,m)},
  \qquad
  \begin{aligned}
    &\prn{\boldsymbol{\epsilon}^{\pm(g,m)}}_{g,g\pm m}
      &\!\!\!\!\!\!&= -\epsilon,\\
    &\prn{\boldsymbol{\epsilon}^{\pm(g,m)}}_{g\pm k,g\pm(k+1)}
      &\!\!\!\!\!\!&= \epsilon
        &\;
        &\forall\, k \in [0,m-1],\\
    &\prn{\boldsymbol{\epsilon}^{\pm(g,m)}}_{g\pm k,g\pm k}
      &\!\!\!\!\!\!&= -\epsilon
        &
        &\forall\, k \in [1,m-1].
  \end{aligned}
\end{equation}
%
But more interestingly for our purposes:
%
\begin{align}
%  \begin{aligned}
\label{eq:shortcutarea}
    \widetilde{\Delta}\potdep _{g,m}\CL &=
      \brk{ \sum_{k=0}^{m-1} \frac{\mu\potdep _{g\pm k,g\pm(k+1)}}{\eq_{g\pm k}}
      - \frac{\mu\potdep _{g,g\pm m}}{\eq_g}}
      + 2rf\potdep  \sum_{k=0}^{m-1} \prn{\eta^+_{g\pm k} - \eta^+_{g\pm(k+1)}} \prn{c_{g\pm k} - c_{g}},
%  \end{aligned}
\end{align}
%
